A free/open-source Kazakh-Tatar machine translation system

### 2013-04-06 paper/paper.tex
more poking on previous work section

Several previous works on making machine translation systems between
Turkic languages exist, although to our knowledge none are publically available.
%FIXME: tr-az, tr-ky
For systems between Turkish and other Turkic languages, there have been systems reported for Turkish-Crimean Tatar \citep{altintas01},
Turkish-Azerbaijani \citep{hamzaoglu93}, Turkish-Tatar \citep{suleymanov08}, and
Turkish-Turkmen \citep{tantug07}, though none of these have been released to a public audience.

A prototype Tatar-Bashkir machine translation system was built by the authors of this paper \citep{tyerswashingtonsalimzyanbattalov12}.
Due to the closeness of the languages, it proved to provide high accuracy
translations, but being a prototype system by design, had relatively low
coverage.

%maybe this should be moved to 'Future plans'
Inside the Apertium project, work on several Turkic languages started to be
carried out, with the ultimate goal of creating independent but compatible
finite-state transducers for each language.
%% The idea is that we started Kazakh-Tatar pair to be able to take into account
%% characteristics of two more distant Turkic languages, avoiding narrowing
%% ourselves to mutually intellegible Tatar and Bashkir, but generate
%% transducers which are still compatible and very coherent. Bashkir part of the
%% pair will be updated to the latest developments of Tatar transducer, which
%% took place while working on Kazakh-Tatar.

>>>>

Within the Apertium project, work on MT systems between Turkic languages has been started (Turkish-Kyrgyz, Azeri-Turkish), but the Kazakh-Tatar system described by the present study is the closest to production-ready of them.  Among these systems is a prototype Tatar-Bashkir machine translation system which was built by the authors of this paper \citep{tyerswashingtonsalimzyanbattalov12}; due to the closeness of these languages, it proved to provide high accuracy in its translations, but being a prototype system by design, had relatively low coverage.

Besides these systems, several previous works on making machine translation systems between Turkic languages exist, although to our knowledge none are publically available.
Some MT systems have been reported that translate between Turkish and other Turkic languages, including Turkish-Crimean Tatar \citep{altintas01},
Turkish-Azerbaijani \citep{hamzaoglu93}, Turkish-Tatar \citep{suleymanov08}, and
Turkish-Turkmen \citep{tantug07}, though none of these have been released to a public audience.

%FIXME: mention tr-az, tr-ky?

%maybe this should be moved to 'Future plans'
Within the Apertium project, work on several Turkic-language MT systems has been started, with the ultimate goal of creating independent but compatible
finite-state transducers for each language.
%% The idea is that we started Kazakh-Tatar pair to be able to take into account
%% characteristics of two more distant Turkic languages, avoiding narrowing
%% ourselves to mutually intellegible Tatar and Bashkir, but generate
%% transducers which are still compatible and very coherent. Bashkir part of the
%% pair will be updated to the latest developments of Tatar transducer, which
%% took place while working on Kazakh-Tatar.
